---
title: "CAT Bond Modeling - Back of the Envelope"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(EnvStats)
library(tidyverse)
library(stats)
library(raster)
library(rgdal)
library(sf)
library(rgeos)
library(ncdf4)
library(rstan)

knitr::opts_chunk$set(echo = TRUE)
setwd("/users/matt/desktop/work/allfed/resilience/modeling")
```

# Parameters

```{r}

# Parameters

######## Modeling ########
n <- 10000 # number of simulations

warming <- c(0.05,1,1.5)

```

# Modules

## Exposure module

**Module purpose**: Define the hazard phenomena: for example, weather events that may impact the food supply. Simulate these events to get a distribution for their frequency and severity.

At this stage, we use historical temperature data to derive a distribution for maximum air temperature during the growing season. Historical data comes from the [FAO](http://www.fao.org/3/y4347e/y4347e0x.htm). Indonesia has several growing seasons, so we treat the distribution as a "bag of growing months." Preliminarily, we treat all months and locations as the same for back-of-the-envelope purposes. We fit a distribution to this figure.

```{r}

# We want to come up with an annual distribution for mean growing season temperatures in Indonesia. For now, we use the last thirty years of data and assume a Gaussian distribution

earliest.year <- 1990

# It's planting season almost every month in various parts of Indonesia
growing.season.months <- c("MAR","APR","MAY","JUN","AUG","SEP","OCT","NOV","DEC")

history <- read.delim("data/temperature/ind.txt", sep="", skip=3) # historical monthly temperature data
history.small <- history %>% subset(YEAR >= earliest.year) %>% dplyr::select(growing.season.months)

#mean.temperatures <- unlist(history.small) # frame to vector of months
mean.temperatures <- rowMeans(history.small) # here averaging across growing season months to get a "mean growing season temperature"

# Fit a distribution to mean temperatures (very simple stan model)
data <- list(temp=mean.temperatures, N=length(mean.temperatures))
temperature.fit <- stan("temp_model.stan", data=data)
posterior <- extract(temperature.fit)

# Fetch parameters - median of the posterior distribution for each
mean_ <- median(posterior$mean_)
sigma_ <- median(posterior$sigma)

# Generate n samples according to model parameters (in this case normal)

### Temperature dist
temperatures <- rnorm(n, mean_, sigma_)

```

## 2) Hazard module

**Module purpose**: Quantify the impact that the events described in the module (1) have on the food supply. Estimate a distribution of food supply shocks over all locations of interest.

Historical crop production data from [FAO](http://www.fao.org/faostat/en/#data/QC/visualize).

```{r}

# source: http://www.fao.org/faostat/en/#data/QC/visualize

fao.all <- read.csv("data/fao/Production_Crops_E_All_Data/Production_Crops_E_All_Data.csv")
fao.indonesia.rice <- fao.all %>% subset(Area == "Indonesia" & Item.Code == 27 & Element.Code == 5510) # historical rice production in Indonesia
indonesia.rice.reshaped <- data.frame(year=colnames(fao.indonesia.rice), value=c(t(fao.indonesia.rice)))
indonesia.rice.years <- indonesia.rice.reshaped %>% mutate(year = as.numeric(str_extract(year, "(?<=Y)[0-9][0-9][0-9][0-9]$")), value=as.numeric(value)) %>% subset(!is.na(year))

# shortcut to projecting the next five years: linear regression over the past thirty to get a trendline, then predict 2.5 years down the road and extract gaussian values around it
trend <- lm(value ~ year, data=indonesia.rice.years)
prediction_2.5 <- predict.lm(trend, data.frame(year=2026))

### Yield dist
yields <- rnorm(n, prediction_2.5, sd(indonesia.rice.years[-5:-1,'value'])) # using the variance over the past five years

```

## 3) Vulnerability module

**Module purpose**: Combine exposure and hazard modules to estimate the potential damage at each location of interest (in this case the whole country).

Here we just assume an effect of temperature on yield based on estimates from [2]: -8.3% plus or minus 1.4% yield loss per degree Kelvin of warming (K = C + 273.15).

```{r}

yield.percentage.change <- (temperatures - mean(temperatures) + 1) * rnorm(n, -0.083, 0.007)
yield.total.change <- yields*yield.percentage.change

# set non-losses to 0.
lost.yields.only <- yield.total.change
lost.yields.only[lost.yields.only > 0] <- 0
lost.yields.only <- abs(lost.yields.only)

```

## 4) Financial module

**Module purpose**: Translate physical damage into monetary cost

As of right now, the [wholesale price](https://www.tridge.com/intelligences/rice/ID) of Indonesian rice is \$0.78/kg, or \$780 / ton. We convert crop production losses into dollar amounts using this figure.

```{r}

cost.per.ton.of.rice <- 780

loss.curve <- lost.yields.only * cost.per.ton.of.rice

# Average annual loss = area under the EP curve = sum product of mean loss and annual likelihood of occurrence - AAL is also known as the "pure premium."
# good reference for AAL: air-worldwide.com/publications/air-currents/2013/Modeling-Fundamentals--What-Is-AAL-/

aal <- mean(loss.curve)

# Here we come up with the exceedance probability (EP) curve: probability of exceeding the given amount in any given year. This can also be expressed as table (probability/loss/return period). It's just another way of visualizing the loss distribution.

loss.amount <- seq(from = 0, to = max(loss.curve), length.out=n)
ep <- sapply(loss.amount, function(x) { return(mean(loss.curve > x)) }, simplify="vector")
plot(loss.amount, ep, type="l") # plot EP curve

# Get coefficient of variation for loss curve: CV = normalized variation
# Variability of loss curve = risk => affects premium
sd.loss <- sd(loss.curve)
cv <- sd.loss/aal


```

# Bond pricing

The United Nations Central Emergency Response Fund ([CERF](https://cerf.un.org/what-we-do/allocation/2020/sector/29/20-RR-SSD-40943/20-RR-WFP-015)) disburses food aid in the event of some severe food catastrophes. Recently, CERF disbursed $15m to feed 90,000 people over an 8-month period in South Sudan, suggesting a cost of ~\$160 per person over the time period.

We suppose here that CERF would want to insure against the equivalent of a roughly 2% loss in rice production in Indonesia relative to this year's production (~1.6 million tons), and assume conservatively that they insure rice at its wholesale market rate.

```{r }

bond.term = 5

#### Policy amount

rice.production.this.year <- 80000000 # rounded
production.loss <- 0.02
policy.amount <- rice.production.this.year * production.loss * cost.per.ton.of.rice

### Bond stuff
bond.face.value <- policy.amount
payout.risk <- mean(loss.curve > bond.face.value)

# Setting annual coupon to expected loss with annual payours
coupon <- payout.risk / bond.term

###### Other (currently unused) ######

expense.ratio = 0.15 # apparently standard
gross.premium = aal / (1 - expense.ratio)

```

# Data sources

A. [Climate research unit - monthly temperature data](https://crudata.uea.ac.uk/cru/data/hrg/cru_ts_4.03/crucy.1905151143.v4.03/new_countries/)

# End notes

1. [Assessing risks of climate variability and climate change for Indonesian rice agriculture](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1876519/)
2. [Plausible rice yield losses under future climate warming](https://www.nature.com/articles/nplants2016202#:~:text=More%20than%2080%25%20(67%20out,%25%20K%E2%88%921%20(Fig.)
3. [Food security, farming and climate change to 2050](https://www.ifpri.org/publication/food-security-farming-and-climate-change-2050)

# Modeling References

1. [Catastrophe Modeling: Guidance for Non-Catastrophe Modellers](https://www.lmalloyds.com/lma/finance/Cat_Modelling_Guidance.aspx)
2. [A Guide to Catastrophe Modeling](https://forms2.rms.com/rs/729-DJX-565/images/rms_guide_catastrophe_modeling_2008.pdf)
3. [Establishing flood damage functions for agricultural crops using estimated inundation depth and flood disaster statistics in data-scarce regions](https://www.jstage.jst.go.jp/article/hrl/11/1/11_12/_article)
4. [Evaluating CAT Bond Pricing Methods](https://pdfs.semanticscholar.org/c2d2/2651f9cc8504a998f8da4ef2a73c4a2d63ef.pdf)
5. [Modeling the Potential Impact of Catastrophic Weather on Crop Insurance Industry Portfolio Losses](https://www.researchgate.net/publication/23521744_Modeling_the_Potential_Impact_of_Catastrophic_Weather_on_Crop_Insurance_Industry_Portfolio_Losses)
6. [Income Elasticities and Global Values of a Statistical Life](https://law.vanderbilt.edu/phd/faculty/w-kip-viscusi/355_Income_Elasticities_and_Global_VSL.pdf) (unused here but included as a reference for later)
7. [Modeling Uncertainty in Climate
Change: A Multi-Model Comparison](https://globalchange.mit.edu/sites/default/files/MITJPSPGC_Rpt290.pdf)
8. [An Evaluation of Decadal Probability Forecasts from State-of-the-Art Climate Models](https://journals.ametsoc.org/jcli/article/26/23/9334/99441/An-Evaluation-of-Decadal-Probability-Forecasts)
  * Features "probability forecasts from ensembles" - using kernel dressing

# Indonesia
* [Rice growing seasons in Indonesia](http://www.fao.org/3/y4347e/y4347e0x.htm)

# Background
* [Financing Catastrophe Risk for Famine Relief Using Early Warning Systems](http://globalagrisk.com/Pubs/2002%20Financing%20Catastrophe%20Risk%20for%20Famine%20Relief%20Using%20Early%20Warning%20jrs.pdf)
* [2004 *Economist* article on potential WFP famine insurance](https://www.economist.com/finance-and-economics/2004/12/09/hedging-against-the-horsemen)

# Aquaculture references
* [The Relative Caloric Prices of Healthy and Unhealthy Foods Differ Systematically across Income Levels and Continents](https://academic.oup.com/jn/article/149/11/2020/5535433)
* [The Future of Food from the Sea](https://oceanpanel.org/sites/default/files/2019-11/19_HLP_BP1%20Paper.pdf)

# Helpful Miscellany
* [Bond valuation on Investopedia](https://www.investopedia.com/terms/b/bond-valuation.asp)
* [A useful blog post](https://understandinguncertainty.org/node/622)
* [Overnight risk-free rates: A User's Guide](https://www.fsb.org/wp-content/uploads/P040619-1.pdf)